---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
```

# Replication analysis for: "Arctic plants can take up inorganic nitrogen year-round"

### Cite this work

Andersen, E.A.S. (2026). Replication analysis for: Arctic plants can take up inorganic nitrogen year-round.

### General information

This code repository is supplement to the publication "Emil Alexander Sherman Andersen<sup>1</sup>, Gesche Blume-Werry<sup>2</sup>, Chenxin Feng (冯晨辛)<sup>1,3</sup>, Friederike Gehrmann<sup>1</sup>, Niki Leblans<sup>1</sup>, Anders Michelsen<sup>4</sup>, Johan Olofsson<sup>2</sup>, Josefine Walz<sup>1</sup>, Ellen Dorrepaal<sup>1</sup>. 2026. Arctic plants can take up inorganic nitrogen year-round. New Phytologist"

DOI: [10.1111/nph.71273](https://doi.org/10.1111/nph.71273)

<small><sup>1</sup>Climate Impacts Research Centre, Department of Ecology and Environmental Sciences, Umeå University, 981 07 Abisko, Sweden  
<sup>2</sup>Department of Ecology and Environmental Sciences, Umeå University, 901 87 Umeå, Sweden  
<sup>3</sup>Faculty of Environmental and Agricultural Sciences, Agricultural University of Iceland, Hvanneyri, IS-311 Borgarness, Iceland  
<sup>4</sup>Terrestrial Ecology Section, Department of Biology, University of Copenhagen, 2100 Copenhagen Ø, Denmark</small>

For questions contact

* Emil A.S. Andersen at ea.sherman@proton.me
* Orcid ID: [0000-0003-3745-5044](https://orcid.org/0000-0003-3745-5044)


### Publication summary:

● Arctic tundra experiences strong climatic seasonality, with cold and long winters, but effective insulation by snow might enable plants and microbes to remain active in winter.  
● We investigated year-round seasonality of plant and microbial inorganic nitrogen (N) uptake, within-plant allocation, and their dependence on snow depth in low-Arctic tundra, using monthly in situ <sup>15</sup>N-pulse-labelling.  
● Plants and microbes took up the <sup>15</sup>N-label throughout the year, likely due to sufficiently deep and early snow that stabilised soil temperatures around 0°C in both study sites. Surprisingly, we found only a few indications for higher plant and microbial <sup>15</sup>N-uptake and (estimated) plant total inorganic N-uptake during the growing season compared to the cold season (i.e., the non-growing season). Instead, N-acquisition strongly varied within the cold season, with clearly higher uptake during the cold deep-winter (November–March) than the warmer spring-winter (April–May/June).  
● Taken together, our results suggest that 78–82% of the annual plant inorganic N-uptake may take place during the long cold season and as much as 50–56% during deep-winter. This demonstrates the importance of nutrient dynamics during the cold season in the Arctic and challenges the general assumption that arctic plants remain dormant during the winter.


### How to use analysis

### Versioning

* Download the code from this repository and the corresponding data from zenodo. Unpack the data (to have two folders: *clean_data* and *raw_data*) and run the "Study I.Rmd" Rproject in Rstudio
* The code was compiled under R version 4.2.3. To access the R package versions used for this analysis use the [renv](https://rstudio.github.io/renv/index.html) package: Open the R project and run in the console:

```{r, results='hide'}
# install.package("renv") # if the renv package is not installed
renv::restore()
```

#### Run the analysis

* All analyses were done in the **15N_analysis.R** script
* For cleaning of data files refer to **Cleaning_script_v2.R**
* For environmental data, cleaning and graph, refer to **Environmental_data.R**
* RAW files can be provided upon request